API design for machine learning software: experiences from the scikit-learn project

### 2013-06-27 paper.tex
copyedit
* Introduce supervised learning, classification and regression
* Explain rbf
* Scikit-learn is a library, not a framework

\title{ API design for machine learning software: experiences from the
scikit-learn project}

@@

In itself, the project is a programming library, i.e. a collection of classes
and functions that users import into Python programs. Using \sklearn therefore
requires basic Python programming knowledge. No command-line interface, let
alone a graphical user interface, is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offer a
\textsc{matlab}-like working environment specifically designed for scientific
use.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section \ref{sec:implementation} briefly describes the implementation.
Section \ref{sec:comparison} then
compares \sklearn to other major projects in terms of API.
Section \ref{sec:future_work} outlines some of the work to be done before
a \sklearn 1.0 release.
We summarize and conclude in section \ref{sec:conclusions}.

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a classification or regression task, the goal is to find a mapping
between input variables $X_1, ..., X_p$, called features, and
some output variable $Y$. A sample is then defined as a pair of
values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
representation of a dataset, a collection of
such samples, is a pair of matrix with numerical values: one for the input
values and one for the output values. Each row of these matrix
corresponds to one sample of the dataset and each column to one variable of
the problem.

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification.
The only requirement is that the \texttt{score} method return a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also exposes a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
y_pred = clf.predict(X_test)
\end{pythoncode}
Note that, in the above example, two distinct hyper-parameter grids are
considered. This is because it only makes sense to search for \texttt{gamma}
when \texttt{kernel} is set to \texttt{"rbf"}.
An SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
so using a single parameter grid would waste computing time
trying out effectively equivalent settings.
Additionally, we see that
\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier:
it delegates the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
\texttt{score} methods to the best estimator
(optionally after re-fitting it on the whole training set).

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). The target audience
of the \sklearn is capable of programming, and therefore we focus on developing
a usable and consistent API, rather than expend effort into creating a GUI. In
addition, running machine learning through a GUI makes reproducibility difficult
(although Weka has developed a GUI for managing such workflows).

Other existing machine learning packages
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
The focus of \sklearn to incorporate the entire package into a single package still
allows for the use of external tools to process input/output, however gives the added
benefit of being able to run these processes within the same development environment.
In addition, this makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

@@

We have discussed the \sklearn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We have shown how a consistent API across the package makes \sklearn
very \textbf{usable} in practice: experimenting with different learning
algorithm is as simple as substituting a new class definition.
Through composition interfaces such as Pipelines, Feature Unions,
and meta-estimators, these simple building blocks lead to an API which is
\textbf{powerful}, and can accomplish a wide variety of learning tasks
within a small amount of easy-to-read code.
Through duck-typing, the consistent API leads to a framework that is
easily \textbf{extensible}, and allows user-defined estimators to be
incorporated into the \sklearn workflow without any explicit object
inheritance.

@@

The authors and contributors acknowledge active support from INRIA. Past and
present sponsors of the project also include Google for funding
scholarships through its Summer of Code program,
the Python Software Foundation and Tinyclues for funding coding sprints.

>>>>

\title{API design for machine learning software: experiences from the
scikit-learn project}

@@

In itself, the project is a programming library, i.e.\ a collection of classes
and functions that users import into Python programs. Using \sklearn therefore
requires basic Python programming knowledge. No command-line interface, let
alone a graphical user interface, is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offer a
\textsc{matlab}-like working environment specifically designed for scientific
use.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the work to be done before
a \sklearn 1.0 release.
We summarize and conclude in section~\ref{sec:conclusions}.

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a supervised learning task, the goal is to find a mapping
from input variables $X_1, \ldots X_p$, called features, to
some output variables $Y$. A sample is then defined as a pair of
values $([x_1, \ldots, x_p]^\mathrm{T}, y)$ of these variables. A widely used
representation of a dataset, a collection of
such samples, is a pair of matrices with numerical values: one for the input
values and one for the output values. Each row of these matrices
corresponds to one sample of the dataset and each column to one variable of
the problem.

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification.
The only requirement is that the \texttt{score} method return a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also expose a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
y_pred = clf.predict(X_test)
\end{pythoncode}
Note that, in the above example, two distinct hyper-parameter grids are
considered. This is because it only makes sense to search for \texttt{gamma}
when \texttt{kernel} is set to a radial basis function kernel \texttt{"rbf"}.
An SVM with a linear kernel accepts a \texttt{gamma} parameter, but ignores it,
so using a single parameter grid would waste computing time
trying out effectively equivalent settings.
Additionally, we see that
\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier:
it delegates the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
\texttt{score} methods to the best estimator
(optionally after re-fitting it on the whole training set).

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). The target audience
of \sklearn is capable of programming, and therefore we focus on developing
a usable and consistent API, rather than expend effort into creating a GUI\@. In
addition, running machine learning through a GUI makes reproducibility difficult
(although Weka has developed a GUI for managing such workflows).

Other existing machine learning packages
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
The focus of \sklearn to incorporate the entire process into a single package still
allows for the use of external tools to process input/output, however gives the added
benefit of being able to run these processes within the same development environment.
In addition, this makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

@@

We have discussed the \sklearn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We have shown how a consistent API across the package makes \sklearn
very \textbf{usable} in practice: experimenting with different learning
algorithm is as simple as substituting a new class definition.
Through composition interfaces such as Pipelines, Feature Unions,
and meta-estimators, these simple building blocks lead to an API which is
\textbf{powerful}, and can accomplish a wide variety of learning tasks
within a small amount of easy-to-read code.
Through duck-typing, the consistent API leads to a library that is
easily \textbf{extensible}, and allows user-defined estimators to be
incorporated into the \sklearn workflow without any explicit object
inheritance.

@@

The authors and contributors acknowledge active support from INRIA\@. Past and
present sponsors of the project also include Google for funding
scholarships through its Summer of Code program,
the Python Software Foundation and Tinyclues for funding coding sprints.
